Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2023-03-24 2_SSL_EES/article/article_supplementary.tex
more edits

The manufacturing process of semiconductor devices can be categorized by the level at which manufacturing process steps are implemented, i.e., either at the wafer level or at the individual chip/package level. While the upstream steps in the manufacturing flow are carried out at the wafer level, the packaging steps are typically carried out at the die/package level. The total manufacturing cost per die is thus the sum of the total costs of all wafer processing steps and all die packaging steps.

@@

\begin{align*}
    P_S &\dots \text{sapphire substrate price} \\
    C_w &\dots \text{wafer processing cost} \\
    C_p &\dots \text{die processing cost}
\end{align*}

@@

\begin{myequation}
        C_w = \sum_i C_i \\
	C_p = \sum_j C_j
\end{myequation}

@@

\begin{myequation}
\label{eqn:cost_wafer}
    C_i \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] =\frac{1}{D}  \frac{1}{y_i}   \bigg\{ \bigg((e*p) + l + m + d +o \bigg)_i \bigg( \frac{t_i}{w_i u_i} \bigg) + \sum_{x} v_x p_x \bigg\}
\end{myequation}

where the index $i$ runs over all wafer processing steps, the index $j$ runs over all die processing steps and the indices $x,y$ run over all materials.

\begin{align*}
        D &\dots \text{functional die per wafer (tested successfully)} \label{def:cost_wafer_first} \\
        y &\dots \text{process step yield} \\
        u &\dots \text{equipment utilization (in relation to theoretical equipment capacity)} \\
        p &\dots \text{power consumption} \\
        e &\dots \text{hourly electricity cost} \\
        m &\dots \text{hourly maintenance cost} \\
        d &\dots \text{hourly depreciation cost}\\
        l &\dots \text{hourly labour cost} \\
        o &\dots \text{hourly overhead cost} \\
        t_i &\dots \text{time per run} \\
        w &\dots \text{wafers per run} \\
        A &\dots \text{wafer area} \\
        v_x &\dots \text{volume of material $x$ per wafer} \\
        p_x &\dots \text{price of material $x$ per volume}\\
\end{align*}

@@

\begin{myequation}
\label{eqn:dpw}
	D=\frac{\pi}{4}  \bigg ( \frac{d-2e}{\sqrt{a}+s/2} \bigg ) ^2 - \frac{\pi}{\sqrt{2}}\frac{d-2e}{(\sqrt{a}+s/2)^2}
\end{myequation}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=5.5cm]{./figures/die-per-wafer.pdf}
    \caption{Number of die per wafer for the different wafer sized used in different years: (DPW) $d(2002)=50$mm$\rightarrow851$DPW to $d(2020)=200$mm$\rightarrow26,838$DPW. Source: own elucidations, based on  \cref{eqn:dpw}.}
    \label{fig:dpw}
\end{figure}

Results are plotted in \cref{fig:dpw}, showing the increase in die per wafer (DPW). \cref{eqn:dpw} then gives us for the cost of a manufacturing step $C_i$ in the wafer processing category

\begin{myequation}
\label{eqn:cost_wafer_full}
\begin{split}
    C_i \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] &= \bigg (  \frac{\pi}{4}  \bigg ( \frac{d-2e}{\sqrt{a}+s/2} \bigg ) ^2 - \frac{\pi}{\sqrt{2}}\frac{d-2e}{(\sqrt{a}+s/2)^2} \bigg )^{-1} \times \\
    &  \frac{1}{y_i}  \bigg\{ \bigg((e*p) + l + m + d +o \bigg)_i \bigg( \frac{t_i}{w_i u_i} \bigg) + \sum_{x} v_x p_x \bigg\}
\end{split}
\end{myequation}

and the cost of a manufacturing step $C_j$ in the packaging category

\begin{myequation}
\label{eqn:cost_die}
    C_j \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] = \frac{1}{y_j}  \bigg\{ \bigg((e*p) + l + m + d + o \bigg)_i  \frac{c_j}{u_j} + \sum_{x} a v_x p_x \bigg\}
\end{myequation}

where $c_j$ is $\text{throughput}^{-1}$. The total cost is thus
\begin{myequation}
\label{eqn:cost_total}
\begin{split}
    C= P_s &+ \sum_i \bigg \{ \frac{1}{D} \frac{1}{y_i} \bigg[ \frac{t_i}{w_i u_i} \bigg((e*p) + l + m + d +o \bigg)_i  + \sum_{x} v_x p_x \bigg] \bigg \} + \\
    & + \sum_j \bigg \{ \frac{1}{y_j} \bigg[ \frac{c_j}{u_j}  \bigg((e*p) + l + m + d + o \bigg)_i + \sum_{x} a v_x p_x \bigg ] \bigg\}
\end{split}
\end{myequation}

Note that in keeping with the categorization introduced by the United States Department of Energy (cf. eg. \cite{doe_ssl_rnd_2019}), certain steps from these two categories are reported separately. In the wafer processing category, the epitaxy step is reported separately due to its complexity and the large share of cost carried. In the wafer processing category, the phosphor step is reported separately.

@@

Devices may be damaged or otherwise rendered unusable during the manufacturing process. The ratio between the number of good devices per step and the number of handled devices per step is known as the yield. Optimizing this yield is critical for reducing manufacturing cost \cite{Kumar2006}. This is because cumulative yield quickly drops as the yield from manufacturing steps with below 100\% yield is multiplied. We must thus consider not only the manufacturing cost per process step, but also the cost including the yield \cite{becker2001use}\cite{becker2001using}. While there are different mathematical approaches to including yield, we follow the definition in \cite{becker2001use}. We write for the yielded cost $C_{Y_i}$ of a step $i$

@@

It can be shown by induction that the general term for a step $i>1$ repeated $n$ times gives an equation of the form:

@@

This cumulative approach to yielded cost is different from the approach taken in the original \textit{LEDCOM} model. It uses what in the literature is described as an \textit{"itemized approach"} to yielded cost \cite{becker2001use}. Using this approach, the yielded cost of a single process step $f$ is described by

@@

\begin{myequation}
\label{eqn:series}
    f_{\times 2} =  \frac{\frac{i+s}{y}+s}{y} \\
    f_{\times n} = \frac{i + s(1+y+y^2+ \dots + y^{n-1})}{y^n}
\end{myequation}

@@

\subsection{Computation of the Contribution of Variables}

@@

\subsection{Comparison with Reported Industry Data}

We note differences between our model and costs reported to the US Department of Energy (DoE) by industry. For instance, the share of the epitaxy step is larger in the DoE model. This can in part be explained by our model relying on state-of-the-art equipment, while industry might not run low-power and mid-power chip production on these, more expensive, reactors. The packaging part of the manufacturing process is increasing in importance, which has been confirmed by researchers and industry reports on wafer-level packaging.

>>>>

The manufacturing process of semiconductor devices can be categorized by the level at which manufacturing process steps are implemented, i.e., either at the wafer level or at the individual chip/package level. The total manufacturing cost per die is thus the sum of the total costs of all wafer processing steps and all die packaging steps.

@@

\begin{align*}
    P_S &\dots \text{sapphire substrate price per die} \\
    C_w &\dots \text{wafer processing cost per die} \\
    C_p &\dots \text{die processing cost}
\end{align*}

@@

\begin{myequation}
        C_w = \sum_i C_i
\end{myequation}
\begin{myequation}
	C_p = \sum_j C_j
\end{myequation}

@@

\begin{myequation}
\label{eqn:cost_wafer}
    C_i \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] =\frac{1}{D}  \frac{1}{y_i}   \bigg\{ \bigg((c_e*p) + c_l + c_m + c_d + c_o \bigg)_i \bigg( \frac{t_i}{w_i u_i} \bigg) + \sum_{x} v_x \c_x \bigg\}
\end{myequation}

where the index $i$ runs over all wafer processing steps, the index $j$ runs over all die processing steps and the index $x$ run over all materials.

\begin{align*}
        D &\dots \text{number of functional (i.e., successfully tested) die per wafer}\\
        y &\dots \text{process step yield} \\
        u &\dots \text{equipment utilization (relative to theoretical equipment capacity)} \\
        p &\dots \text{power consumption} \\
        c_e &\dots \text{hourly electricity cost} \\
        c_m &\dots \text{hourly maintenance cost} \\
        c_d &\dots \text{hourly depreciation cost}\\
        c_l &\dots \text{hourly labour cost} \\
        c_o &\dots \text{hourly overhead cost} \\
        t_i &\dots \text{time per run} \\
        w &\dots \text{wafers per run} \\
        A &\dots \text{wafer area} \\
        v_x &\dots \text{volume of material $x$ per wafer} \\
        c_x &\dots \text{cost of material $x$ per volume}\\
\end{align*}

@@

\begin{myequation}
\label{eqn:dpw}
	DPW = \frac{\pi}{4}  \bigg ( \frac{d-2e}{\sqrt{a}+s/2} \bigg ) ^2 - \frac{\pi}{\sqrt{2}}\frac{d-2e}{(\sqrt{a}+s/2)^2}
\end{myequation}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=5.5cm]{./figures/die-per-wafer.pdf}
    \caption{Number of die per wafer for the different wafer sizes used in different years: (DPW) $d(2002)=50$mm$\rightarrow851$DPW, $d(2002)=50$mm$\rightarrow5370$DPW, $d(2020)=100$mm$, \rightarrow26,838$DPW. Source: own estimates, based on \cref{eqn:dpw} and the dominant wafer size of the considered year, according to \cref{fig:wafers}.}
    \label{fig:dpw}
\end{figure}

Results are plotted in \cref{fig:dpw}, showing the increase in die per wafer (DPW) over time. \cref{eqn:dpw} then gives us for the cost of a manufacturing step $C_i$ in the wafer processing category
\begin{myequation}
\label{eqn:cost_wafer_full}
\begin{split}
    C_i \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] &= \bigg (  \frac{\pi}{4}  \bigg ( \frac{d-2e}{\sqrt{a}+s/2} \bigg ) ^2 - \frac{\pi}{\sqrt{2}}\frac{d-2e}{(\sqrt{a}+s/2)^2} \bigg )^{-1} \times \\
    &  \frac{1}{y_i}  \bigg\{ \bigg((c_e*p) + c_l + c_m + c_d + c_o \bigg)_i \bigg( \frac{t_i}{w_i u_i} \bigg) + \sum_{x} v_x \c_x \bigg\}
\end{split}
\end{myequation}
and the cost of a manufacturing step $C_j$ in the packaging category
\begin{myequation}
\label{eqn:cost_die}
    C_j \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] = \frac{1}{y_j}  \bigg\{ \bigg((c_e*p) + c_l + c_m + c_d + c_o \bigg)_i  \frac{c_j}{u_j} + \sum_{x} a v_x c_x \bigg\}
\end{myequation}
where $c_j$ is $\text{throughput}^{-1}$. The total cost is thus
\begin{myequation}
\label{eqn:cost_total}
\begin{split}
    C= P_s &+ \sum_i \bigg \{ \frac{1}{D} \frac{1}{y_i} \bigg[ \frac{t_i}{w_i u_i} \bigg((e*p) + l + m + d +o \bigg)_i  + \sum_{x} v_x p_x \bigg] \bigg \} + \\
    & + \sum_j \bigg \{ \frac{1}{y_j} \bigg[ \frac{c_j}{u_j}  \bigg((e*p) + l + m + d + o \bigg)_i + \sum_{x} a v_x p_x \bigg ] \bigg\}
\end{split}
\end{myequation}
Note that in keeping with the categorization introduced by the United States Department of Energy (cf. eg. \cite{doe_ssl_rnd_2019}), certain steps from these two categories are reported separately. In the wafer processing category, the epitaxy step is reported separately due to its complexity and the large share of cost carried. In the wafer processing category, the phosphor step is reported separately.

@@

Devices may be damaged or otherwise rendered unusable during the manufacturing process. The ratio between the number of good devices per step and the number of handled devices per step is known as the yield. Optimizing this yield is critical for reducing manufacturing cost \cite{Kumar2006}. This is because cumulative yield quickly drops as the yield from manufacturing steps with below 100\% yield is multiplied. We must thus consider not only the manufacturing cost per process step, but also the cost including the yield \cite{becker2001use}\cite{becker2001using}. While there are different mathematical approaches to including yield, we follow the definition in \cite{becker2001use}. We write for the yielded cost $C_{Y_i}$ of a step $i$ with associated cost (before considering yield) $C_1$ and yield $Y_1$:

@@

It can be shown by induction that the general form of a term for a step $i>1$ repeated $n$ times can be expressed as:

@@

This cumulative approach to yielded cost is different from the approach taken in the original \textit{LEDCOM} model. It uses what in the literature is described as an \textit{"itemized approach"} to yielded cost \cite{becker2001use}. In this approach, the yielded cost of a single process step $f$ is described as

@@

\begin{myequation}
    f_{\times 2} =  \frac{\frac{i+s}{y}+s}{y}
\end{myequation}
\begin{myequation}
    f_{\times n} = \frac{i + s(1+y+y^2+ \dots + y^{n-1})}{y^n}
\label{eqn:series}
\end{myequation}

@@

\subsection{Comparison with Reported Industry Data and DOE Projections}

We conducted a comparison of the LED manufacturing cost structure produced by our cost model with previously reported US DOE calculations and projections based on the LEDCOM model and industry data provided as part of industry round table discussions (see \cref{fig:costmodel_calibration}). We note some differences between the results of our model and the cost structure reported or projected by the DOE. For instance, the share of the epitaxy step is consistently larger in the DOE data. This can in part be explained by our model relying on state-of-the-art equipment at a virtual US-based manufacturing location, while industry might not run low-power and mid-power chip production on these, more expensive, reactors. In addition, the manufacturing lines of the majority of manufacturers is located in Asia. We also note that the share of the substrate price in the DOE data is much larger than in our model, which can in part be explained by the overestimation of the actual price of sapphire wafers in earlier projections for 2015-2020. Finally, the relative importance of the packaging part of the manufacturing process is very similar in our model and in the DOE results in 2012. However, in the DOE projections it significantly decreases by 2020, while in our model it retains and even increases its share of the total cost. This trend has been independently confirmed by researchers and industry reports on wafer-level packaging \cite{Lee2011WPL}\cite{Xie2013}\cite{ledsmag2017WLP}, showing better performance of our cost model compared to earlier DOE model projections

\section{Computation of Contribution of Individual Variables}

@@
